---
title: "R Notebook"
output: html_notebook
---

# Getting Started

This class will work primarily through notebooks. Notebooks are a super handy way of interleaving code and text. This allows you to experiment and write out your ideas while working through code.

This is a code block. The first bit of code is to load some `libraries` into R. Libraries are just collections of functions that are pre-defined for you (so you don't have to code things yourself.)

Run the code block by
```{r}

list_of_packages <- c("tidyverse", "readxl", "haven")

lapply(list_of_packages, require, character.only = TRUE)

```

If you get an error message, that means that you need to install the packages. This only has to be done once.

```{r}
install.packages(list_of_packages)
```


Add a new code chunk by clicking the Insert button on the toolbar, then selecting R or by pressing Ctrl+Alt+I (Windows) or Cmd+Option+I (Mac).

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the Preview button or press Ctrl+Shift+K (Windows) or Cmd+Shift+K (Mac) to preview the HTML file).

Now we can start!