---
title: "Modify and Summarise"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Manipulate, Modify, Summarise

When it comes to data analysis we will often need to manipulate our data to highlight our interest. `dplyr` provides some amazing tools for doing this manipulation. Let's run through a few examples

## Back with our Gapminder

Say I'm a researcher that is only interested in Africa. I'm not really interested in look at any other continents. I would like to make a reduced data set to just see Africa.

```{r}
gapminder %>% 
  filter(continent == "Africa")
```

A couple of moves are occuring here. The first of which is use of the pipe ` %>% ` . This is a special operator that allows you to chain commands together. However, do not feel compelled to use it right now. The equivalent code of the above would be:

```{r}
filter(gapminder, continent == "Africa")
```

Perhap I'm only interested in data before 1977. I can filter on that too!

```{r}
gapminder %>% 
  filter(continent == "Africa", year <= 1977)
```

Ok, great! I can now assign this to a new object in R that just has these data.

```{r}
df_africa <- gapminder %>% 
  filter(continent == "Africa", year <= 1977)
```

## Adding new columns

Say I want to add a new metric. Life Expectancy divided by the Per Capita GDP. This can be accomplished with the `mutate` function.

```{r}
df_africa %>% 
  mutate(life_exp_gpd = gdpPercap/ lifeExp)
```

Let's save it now!

```{r}
df_africa_2 <-df_africa %>% 
  mutate(life_exp_gpd = gdpPercap/ lifeExp)
```


## Summarise it

With all of this great work, we now want to see what the average is per country within Africa for our new metric. We can do this summary first by `group_by` which will ensure that we get a summary value by the grouping variable. We will then supply a `summarise` argument to give us our summary value.

```{r}
df_africa_2 %>% 
  group_by(country) %>% 
  summarise(mean_life_exp_gpd = mean(life_exp_gpd))
```

And I'm really only interested in the bottom ten countries.  So let's use another function `top_n` to filter four us. We can also supply the `arrange` that will sort by our specification (default is ascending).

```{r}
df_africa_2 %>% 
  group_by(country) %>% 
  summarise(mean_life_exp_gpd = mean(life_exp_gpd)) %>% 
  top_n(-10, wt = mean_life_exp_gpd) %>% 
  arrange(mean_life_exp_gpd)
```

