---
title: "LAB1"
output: html_notebook
---

Now it's your turn. We will be using the `mpg` data set that is pre-loaded in R. You can explore it with

```{r}
library(tidyverse)
```


```{r}
mpg
```

## Question 1
I would like a graph that plots `cty` milleage (x axis) versus `hwy` mileage on with a line. Hint: use `geom_line()`

```{r}
ggplot(data = mpg, aes(cty, hwy))+
  geom_point()
```

##Question 1A

Now color the the dots by the discplacement. You will need to supply a color argument in the `aes`, and use `as.factor(cyl)`.

```{r}
ggplot(data = mpg, aes(cty, hwy, color = as.factor(cyl)))+
  geom_point()
```


## Question 2
Now I would like a histogram that of the `hwy` variable. Hint use `geom_histogram`

```{r}
ggplot(data = mpg, aes(hwy))+
  geom_histogram()
```

## Question 3
Now I would like to see a boxplot of the different drv trains versus the engine displacement. Hint use `geom_boxplot`

```{r}
ggplot(mpg, aes(drv, displ))+
  geom_boxplot()
```

# Question 4
Make a new column that is the hwy mileage (hwy) divided by the engine discplacement (displ). Call this new value `miles_per_l`. Save the new data set as `mpg_2`

```{r}
mpg_2 <- mpg %>% 
  mutate(miles_per_l = hwy/displ)
```

# Question 5
Now group by the car `cyl` type and return a table with the averahe `miles_per_l`.

```{r}
mpg_2 %>% 
  group_by(cyl) %>% 
  summarise(mean_m_p_l = mean(miles_per_l))
```

